---
title: "brs: Bayesian Rule Set"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{brs-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**R** package for *Bayesian Rule Set: A Quantitative Alternative to Qualitative Comparative Analysis.*

**References**: 
 
 - Chiu, Albert and Yiqing Xu. "Bayesian Rule Set: A Quantitative Alternative to Qualitative Comparative Analysis." *The Journal of Politics*, forthcoming. 
 - Wang, Tong, Cynthia Rudin, Finale Doshi-Velez, Yimin Liu, Erica Klampfl, and Perry MacNeille. "A bayesian framework for learning rule sets for interpretable classification." *The Journal of Machine Learning Research** 18, no. 1 (2017): 2357-2393.
 
**R** source files can be found on [GitHub](https://github.com/albert-chiu/brs). **R** code used in this demonstration can also be downloaded from [GitHub](https://github.com/albert-chiu/brs/tree/main/vignettes).


```{r setup}
library(brs)

# create conda environment
reticulate::conda_create("BRS_conda2")  

# install python and packages to environment
reticulate::conda_install(envname = "BRS_conda3",
  conda="~/miniforge3/condabin/conda",  # check default sequence of locations for conda
  #channel = "conda-forge/osx-arm64",
  packages = c("numpy", "pandas", "scikit-learn", "scipy"))
  #python_version = "3.6")  # python version to use
    # note: replication code for paper uses 3.6.1, 
    # but version names have since been truncated
reticulate::use_condaenv(condaenv="BRS_conda3")  ## use this conda environment for BRS

#reticulate::conda_remove("BRS_conda")

set.seed(123)
data("lipset_df", "lipset_Y")
out_lipset <- brs::BRS(df = lipset_df, Y = lipset_Y, 
                       maxLen=3L, trainProp=1)
brs::plot_bar(df = lipset_df, fit = out_lipset, featureLabels = fdf, maxLen=3)

```
