---
title: "brs: Bayesian Rule Set"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{brs-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**R** package for *Bayesian Rule Set: A Quantitative Alternative to Qualitative Comparative Analysis.*

**References**: 
 
 - Chiu, Albert and Yiqing Xu. "Bayesian Rule Set: A Quantitative Alternative to Qualitative Comparative Analysis." *The Journal of Politics* 85, no. 1 (2023):. 
 - Wang, Tong, Cynthia Rudin, Finale Doshi-Velez, Yimin Liu, Erica Klampfl, and Perry MacNeille. "A bayesian framework for learning rule sets for interpretable classification." *The Journal of Machine Learning Research* 18, no. 1 (2017): 2357-2393.
 
**R** source files can be found on [GitHub](https://github.com/albert-chiu/brs). **R** code used in this demonstration can also be downloaded from [GitHub](https://github.com/albert-chiu/brs/tree/main/vignettes).


```{r setup, cache=T}
# create conda environment
#reticulate::conda_create("BRS_conda2")  

# install python and packages to environment
reticulate::conda_install(envname = "BRS_conda3",
  conda="~/miniforge3/condabin/conda",  # check default sequence of locations for conda
  #channel = "conda-forge/osx-arm64",
  packages = c("numpy", "pandas", "scikit-learn", "scipy"))
  #python_version = "3.6")  # python version to use
    # note: replication code for paper uses 3.6.1, 
    # but version names have since been truncated
reticulate::use_condaenv(condaenv="BRS_conda3")  ## use this conda environment for BRS

#reticulate::conda_remove("BRS_conda")

# brs
library(brs)
set.seed(123)
data("lipset_df", "lipset_Y")
out_lipset <- brs::BRS(df = lipset_df, Y = lipset_Y, 
                       maxLen=3L, trainProp=.7, bootstrap = T, reps=10L)
```

```{r barplot}
# barplot
fdf <- cbind(colnames(lipset_df),
             c("Wealth (high)", "Wealth (med)", "Wealth (low)",
               "Urbanization (high)", "Urbanization (low)", 
               "Education (high)", "Education (low)", 
               "Industrialization (high)", "Industrialization (low)"))
oppmat <- matrix(c(0,1), ncol=2)
oppind <- list(unique(unlist(lapply(colnames(lipset_df), function(x) strsplit(x, "_")[[1]][[1]])))[2:4])

lipset_bar <- brs::plot_bar(df = lipset_df, Y=lipset_Y, fit = out_lipset, featureLabels = fdf,
              maxLen=3, minProp = .05, topRules=5, 
                           simplify = T, oppmat=oppmat, oppind=oppind, and =" & ",
                           plotBuffer = c(.25, 0, .4), boot_rep = 10L,
                           titleSize=8, rule_text_size = 8, number_size = 8)
```

```{r print-barplot, fig.dim = c(7, 4)}
print(lipset_bar)
```
